---
title: "Bivariate Chen Distribution: Properties, Estimation, and Applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate Chen Distribution: Properties, Estimation, and Applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(BvChen)
```

The bivariate Chen distribution is built from three independent Chen
(2000) components $X_i \sim Ch(\alpha_i,\beta)$ via
$Z_1=\min(X_0,X_1)$ and $Z_2=\min(X_0,X_2)$. The margins are
$Z_1\sim Ch(\alpha_1+\alpha_3,\beta)$ and
$Z_2\sim Ch(\alpha_2+\alpha_3,\beta)$, and the law carries a singular
component on the diagonal $z_1=z_2$ with mass
$\alpha_3/(\alpha_1+\alpha_2+\alpha_3)$.

## Simulating data

```{r}
set.seed(2024)
sim <- rbvch(400, 0.8, 1.2, 0.5, 1.5)
head(sim)
mean(sim[, "z1"] == sim[, "z2"])   # about 0.5 / 2.5 = 0.2
```

## Fitting with the EM algorithm

```{r}
fit <- fitbvch(sim[, "z1"], sim[, "z2"])
fit
coef(fit)
```

## Profile likelihood for beta

```{r, fig.width=6, fig.height=4}
pb <- profile_beta(sim[, "z1"], sim[, "z2"])
pb$beta_hat
with(pb$grid, plot(beta, prof_loglik, type = "l",
                   xlab = expression(beta), ylab = "profile log-likelihood"))
abline(v = pb$beta_hat, lty = 2, col = "red")
```

## Goodness of fit

```{r}
cf <- coef(fit)
gofbvch(sim[, "z1"], sim[, "z2"], cf["alpha1"], cf["alpha2"],
        cf["alpha3"], cf["beta"])
```

```{r, fig.width=7}
qqbvch(sim[, "z1"], sim[, "z2"], cf["alpha1"], cf["alpha2"],
       cf["alpha3"], cf["beta"])
```
